---
title: "Project Code"
author: "Jill Reiner"
date: "6/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggbeeswarm)
library(ggridges)
library(forcats)
nwsl <- read.csv("http://www.stat.cmu.edu/cmsac/sure/materials/data/eda_projects/nwsl_season_stats.csv")
```

```{r}
playoff_teams <- nwsl %>%
  filter(team_id %in% c("NC", "CHI", "POR", "SEA") & season == 2019 |
         (team_id %in% c("NC", "CHI", "POR", "SEA") & season == 2018) |
         (team_id %in% c("NC", "CHI", "POR", "ORL") & season == 2017)) %>%
  mutate(goals_per_match = goals / matches) %>%
  group_by(team_id, season) %>%
  summarize(avg_goals_per_match = sum(goals_per_match))
```

# Hypothesis: Does outscoring teams in the regular season give you an advantage in the postseason? 

# We need to reword our hypothesis here to match what the plot is telling us...we've tried other plots but want to stick to this one, just need to work on the hypothesis.

# Over the past three NWSL seasons, championship teams have been scoring more goals per match in the regular season.

# Would it be helpful to just look at 2018 and 2019? Would it be better to look at all teams instead of just playoff teams? We just need some guidance on this one.

```{r}
playoff_teams %>%
  ggplot(aes(x = team_id, fill = team_id)) +
  geom_bar(aes(y = avg_goals_per_match),
           stat = "identity") +
  scale_fill_manual("Teams", values = c("CHI" = "lightblue", 
                                         "NC" = "darkblue", 
                                         "POR" = "darkred",
                                         "SEA" = "gold",
                                         "ORL" = "purple")) + 
  labs(x = "Team", y = "Average Goals per Match",
       title = "Regular Season Average Goals per Match",
       subtitle = "NWSL Playoff Teams from 2017-2019") +
  theme_bw() +
  facet_wrap(~ season, ncol = 3)
```

```{r}
last_three_champs_full <- nwsl %>%
  filter((team_id == "NC" & season == 2019) | (team_id == "NC" & season == 2018) | (team_id == "POR" & season == 2017),
         goals > 0) %>%
  mutate(g_permatch = goals / matches)
```

```{r}
last_three_champs_summarized <- last_three_champs_full %>%
  group_by(season, team_id) %>%
  summarise(regseason_goals_per_match = sum(g_permatch))
```

# How does a team's goal scoring ability in the regular season reflect in the postseason?

```{r}
last_three_champs_summarized %>%
  ggplot(aes(x = season, fill = team_id)) +
  geom_bar(aes(y = regseason_goals_per_match),
           stat = "identity") +
  scale_fill_manual("Teams", values = c( "NC" = "darkblue", 
                                         "POR" = "darkred")) + 
  labs(x = "Season", y = "Regular Season Goals per Match",
       title = "Regular Season Goals per Match",
       subtitle = "Last Three NWSL Championship Winning Teams") +
  theme_bw()
```

# Pick between this bar plot and first, can't have 2. We're leaning towards the 1st one, but it's still TBD.

```{r}
team_distribution_2019_nwsl3 <- nwsl %>%
  filter(pos %in% c("FW,MF", "DF,MF", "FW", "MF"),
         season == 2019,
         goals > 0)
```

# Hypothesis: Does having the league leading scorer automatically win you championships?
```{r}
team_distribution_2019_nwsl3 %>%
  ggplot(aes(x = goals,
             y = team_id)) +
  geom_density_ridges() +
  geom_rug(alpha = .03) + 
  labs(x = "Goals", y = "Team",
       title = "Distribution of Goal Scorers per Team",
       subtitle = "2019 NWSL Regular Season") +
  theme_bw()
```

```{r}
team_distribution_2019_nwsl3 %>%
  ggplot(aes(x = goals,
             y = team_id)) +
  geom_boxplot() + 
  labs(x = "Goals", y = "Team",
       title = "Distribution of Goal Scorers per Team",
       subtitle = "2019 NWSL Regular Season") +
  coord_flip() + 
  theme_bw()
```

# Same idea as previous plot, just displayed differently

# NC won championship in 2019 and they had a lot of different goal scorers with pretty good goal totals (as opposed to say, CHI, who had one really good goal scorer but the rest of their players were meh), so this may indicate that it is better to have more depth in terms of goal scoring and it doesn't always help the team as a whole to have the leading scorer on their team. You don't need to have the top goal scorer on your team to win championships. Team work makes the dream work!



